---
title: "CDI_analysis"
output: html_document
---
#A breakdown of the CDM First Word Survey by CDI classification of the word. Some words were omitted due to not being translatable, and some of the words were classified in an ad-hoc manner

---
#Preliminaries 
```{r}
library(plyr)
library(ggplot2)
library(reshape2)
theme_set(theme_bw())
df = read.csv("data/CDM_CDIs.csv") 

```

---
#Freqs of First words - count
```{r}
freqs <- ddply(df, .(Word), summarise, count=length(Word))
freqs$Word <- factor(freqs$Word, levels=with(freqs, Word[order(count, Word, decreasing = TRUE)]))
q <- qplot(Word, count, geom="bar", stat="identity", 
      data=subset(freqs, count>5))
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5))
```

#Freqs of first words props 
```{r}
freqs <- ddply(df, .(Word), summarise, count=length(Word))
freqs$Word <- factor(freqs$Word, levels=with(freqs, Word[order(count, Word, decreasing = TRUE)]))
freqs$prop <- freqs$count/sum(freqs$count)
q <- qplot(Word, prop, geom="bar", stat="identity", 
      data=subset(freqs, count>5))
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5)) + 
  ylab("Proportion of Total Words")
```
---
#Prop of first word by gender 
```{r}
freqs <- ddply(df, .(Word,gender), summarise, count=length(Word))
freqs <- ddply(freqs, .(Word), mutate, total.count = sum(count))
freqs$Word <- factor(freqs$Word, 
                     levels=unique(with(freqs, 
                                 Word[order(total.count, 
                                            Word, decreasing = TRUE)])))
freqs$prop <- freqs$count/sum(freqs$count)
q <- qplot(Word, prop, fill=gender, 
           position="dodge",
           geom="bar",
           stat="identity",
           data=subset(freqs, total.count>5 & gender!=""))
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5)) + 
  ylab("Proportion of Total Words")

```
---

#Freqs of CDIs - Count
```{r}
freqs <- ddply(df, .(CDI_category), summarise, count=length(CDI_category))
freqs$CDI_category <- factor(freqs$CDI_category, 
                     levels=with(freqs, 
                                 CDI_category[order(count, CDI_category, decreasing = TRUE)]))
q <- qplot(CDI_category, count, geom="bar", stat="identity", 
      data=subset(freqs))
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust = .5)) + 
  ylab("Count of Total Words")

               
```
--- 
#Proportions of CDIS 
```{r}
freqs <- ddply(df, .(CDI_category), summarise, count=length(CDI_category))
freqs$CDI_category <- factor(freqs$CDI_category, 
                     levels=with(freqs, 
                                 CDI_category[order(count, CDI_category, decreasing = TRUE)]))
freqs$prop <- freqs$count/sum(freqs$count)

quartz()
qplot(CDI_category, prop, geom = "bar", stat="identity", data=subset(freqs, count>1))+theme_bw() + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5)) +
  ylab("Proportion of Total Words") + xlab("CDI Category")


```

--- 
#Proportions with Ages 
```{r}
freqs <- ddply(df, .(CDI_category, Age_utterance), summarise, count=length(CDI_category))
freqs <- ddply(freqs, .(CDI_category), mutate, total.count=sum(count))
freqs <- ddply(freqs, .(Age_utterance), mutate, prop = count/sum(count))
freqs$prop <- freqs$count/sum(freqs$count)

freqs$CDI_category <- factor(freqs$CDI_category, 
                             levels=unique(with(freqs, 
                                                CDI_category
                                                [order(total.count, CDI_category, decreasing = TRUE)])))

q <- qplot(CDI_category, prop, fill=Age_utterance, 
      position="dodge", geom="bar", stat="identity", data=subset(freqs, total.count>1 & 
                                                                   Age_utterance!="" &
                                                                   Age_utterance!="dontremember"))

q + theme(axis.text.x = element_text(angle=90, hjust = 1))

```
---
#Proportions by gender 
```{r}
quartz()
freqs <- ddply(df, .(CDI_category, gender), summarise, count=length(CDI_category))
freqs <- ddply(freqs, .(CDI_category), mutate, total.count=sum(count))
freqs <- ddply(freqs, .(gender), mutate, prop = count/sum(count))
freqs$prop <- freqs$count/sum(freqs$count)

freqs$CDI_category <- factor(freqs$CDI_category, 
                             levels=unique(with(freqs, 
                                                CDI_category
                                                [order(total.count, CDI_category, decreasing = TRUE)])))

q <- qplot(CDI_category, prop, fill=gender, 
      position="dodge", geom="bar", stat="identity", data=subset(freqs, total.count>1 & gender!=""))
                                                                  
q + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5)) + 
  ylab("Proportion of Total Words") + 
  xlab("CDI Category")

```

----
Age by gender

```{r}
quartz()
mss <- aggregate(Word ~ Age_utterance + gender, df, length)
ms <- ddply(mss, .(gender), function(x) {
            x$prop = x$Word / sum(x$Word)
            return(x) })
            
qplot(Age_utterance, prop, fill=gender, geom="bar", stat="identity",
      position="dodge", 
      data=subset(ms,gender!="" & 
                    Age_utterance!="" & 
                    Age_utterance!="dontremember")) + 
  ylab("Proportion of Children") + 
  xlab("Age at time of first word")

```


--- 
#Age Split - 10m against everyone else 
```{r}
quartz()

df$Age <- factor(c("< 10m","> 10m")[(1 - as.numeric(df$Age_utterance == "10mOrBefore"))+1])
#breaking up each age group individually, just in case we want to do something with it later 
df$eleven <- df$Age_utterance == "11m"
df$twelve <- df$Age_utterance == "12m"
df$thirteen <- df$Age_utterance == "13m"
df$fourteen <- df$Age_utterance == "14mOrLater"

freqs <- ddply(df, .(CDI_category,Age), summarise, count=length(CDI_category))
freqs <- ddply(freqs, .(CDI_category), mutate, total.count = sum(count))
freqs <- ddply(freqs, .(Age), mutate, prop = count / sum(count))

freqs$CDI_category <- factor(freqs$CDI_category,
                             levels=unique(with(freqs, 
                                                CDI_category[order(total.count, CDI_category, decreasing = TRUE)])))
freqs$prop <- freqs$count/sum(freqs$count)


qplot(CDI_category, prop, fill=Age, 
      position="dodge",
      geom="bar", stat="identity", 
      data=subset(freqs,total.count>1)) + theme(axis.text.x = element_text(angle=90, hjust = 1, vjust=.5))+ 
  ylab("Proportion of Total Words") + 
  xlab("CDI Category")

```

--- 
do people report ages differently if their kids are older? 

```{r}
quartz()
df$Age_current <- factor(df$Age_current, levels=c("0to3y","4to6y","7to12y","13plusy"))
df$Age_current <- revalue(df$Age_current, c("0to3y"="0 - 3 years", 
                                            "4to6y"="4 - 6 years", 
                                            "7to12y"="7 - 12 years",
                                            "13plusy"="13+ years"))


mss <- aggregate(Word ~ Age_utterance + Age_current, df, length)
ms <- ddply(mss, .(Age_current), function(x) {
            x$prop = x$Word / sum(x$Word)
            return(x) })

qplot(Age_utterance, prop, fill=Age_current, geom="bar", stat="identity",
      position="dodge", 
      data=subset(ms,Age_current!="" & 
                    Age_utterance!="" & 
                    Age_utterance!="dontremember")) + 
  ylab("Proportion of Children") + 
  xlab("Age at time of first word")
```